Risk Perceptions for Wearable Devices

### 2015-03-04 tex-inputs/intro.tex
abstract redo, intro tweaks, and much progress on data type results section

Along with these benefits, wearable devices bring with them new potential privacy and security risks which expose users' activities without their awareness or consent. Fitbit allowed sex to be tracked as exercise while fitness profiles were public by default~\cite{Fitbit}, resulting in the inadvertent disclosure of sensitive information. Public discomfort toward facial recognition prevented the capability from being deployed to Google Glass \cite{GlassDetection}. Google Glass, the iconic wearable of its time, has since disappeared \cite{13_google_2015}. Most suspect that the reason for its disappearance was because it was the ``straw the broke the back of the privacy camel's back'' \cite{14_dvorak_2014}. Some Glass wearers faced assault \cite{1_russell_2014, 15_mashable_2014, 16_gross_2014}, while bystanders felt uncomfortable personal details, conversations, and photos possibly recorded.

@@

Human attention is a finite resource \cite{bohme2011security}. Given the additional capabilities of wearable devices and their increasing popularity, we need to study user concerns to warn users only about situations they are likely to care about. Informed, select notifications make for a better user experience and prevents habituation to such notifications while avoid scandalous breaches of privacy.

@@

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item Comparisons of users' perceptions of a range of privacy and security risks of wearables. We found that users care much more about the type of data than the recipient of the data.
\item Insight into how users feel about various data recipients. We observed that users make less distinction between sharing data with friends, co-workers, and the general public, comparatively to sharing with an application's servers.
\item Rankings of how data-collection capabilities of wearable devices compared to more familiar technologies.  Most saw new capabilities as benign, but we suspect that this may be due to a lack of exposure to these newer technologies.
\item A taxonomy and report of users' self-reported top concerns for wearable devices. Privacy is by far the top concern, along with security, health risks, financial risk, and social impact. 
\end{itemize}

>>>>

Along with these benefits, wearable devices bring new potential privacy and security risks which expose users' activities without their awareness or consent. Fitbit allowed sex to be tracked as exercise while fitness profiles were public by default~\cite{Fitbit}, resulting in the inadvertent disclosure of sensitive information. Public discomfort toward facial recognition prevented the capability from being deployed to Google Glass \cite{GlassDetection}. Google Glass, the iconic wearable of its time, has since disappeared \cite{13_google_2015}. Most suspect that the reason for its disappearance was because it was the ``straw the broke the back of the privacy camel's back'' \cite{14_dvorak_2014}. Some Glass wearers faced assault \cite{1_russell_2014, 15_mashable_2014, 16_gross_2014}, while bystanders felt uncomfortable personal details, conversations, and photos possibly recorded.

@@

With the additional capabilities of wearable devices and their increasing popularity, people have expressed interest in being notified before data capture \cite{denning2014situ}, but human attention is a finite resource \cite{bohme2011security}. Therefore, user concerns should be investigated to to warn users only about situations they are likely to care about. Informed, select notifications make for a better user experience and prevents habituation to such notifications while avoiding scandalous breaches of privacy.

@@

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item Comparisons of users' perceptions of a range of privacy and security risks of wearables. We found that users care much more about the type of data than the recipient of the data.
\item Insight into how users feel about various data recipients. We observed that users make less distinction between sharing data with friends, co-workers, and the general public, comparatively to sharing with an application's servers.
\item A taxonomy and report of users' self-reported top concerns for wearable devices. Privacy is by far the top concern, along with security, health risks, financial risk, and social impact. 
\item Rankings of how data-collection capabilities of wearable devices compared to more familiar technologies.  Most saw new capabilities as benign, but we suspect that this may be due to a lack of exposure to these newer technologies.
\end{itemize}

### 2015-03-04 tex-inputs/results-recipient.tex
abstract redo, intro tweaks, and much progress on data type results section

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| r | l | r | l |c |}
\hline
Rank & Recipient & VUR & sigma & Distribution \\
\hline
1 & Work Contacts & 75.16\% & 0.938 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/recipient4/recipient_work} \\
2 & Public & 72.41\% & 0.981 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/recipient4/recipient_pub}  \\
3 & Friends & 69.47\% & 1.019 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/recipient4/recipient_friends}\\
4 & App's Server & 42.28\% & 1.150 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/recipient4/recipient_app}\\
\hline
\end{tabular}
\caption{The 10 most and least upsetting data types, across all recipients.}
\label{recipient-table}
\end{center}
\end{table}

%\begin{table}[t]
%\begin{center}
%\begin{tabular}{|l|r|r|r|r|}
%\hline
%Recipients	& $\chi^2$ & p-value 	& n & $\phi$ \\
%\hline
%Work-App	& 565.910 & <0.0001 & 5,083 & 0.111\\
%Public-App	& 481.776 & <0.0001 & 5,1988& 0.093\\
%Friends-App & 381.653 & <0.0001 & 5,096 & 0.075\\
%Friends-Work & 20.39 & <0.0001 & 5,037 & 0.004\\
%Friends-Public & 5.41 & <0.0200 & 5,142 & 0.001\\
%Work-Public&  5.00 & <0.0253 & 5,129	& 0.001\\
%\hline
%\end{tabular}
%\caption{Results of a chi-square test to examine VUR based on data recipient, across all data points.}
%\label{recipient}
%\end{center}
%\end{table}

>>>>

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| r | l | r | l |c |}
\hline
Rank & Recipient & VUR & sigma & Distribution \\
\hline
1 & Work Contacts & 75.16\% & 0.94 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/recipient4/recipient_work} \\
2 & Public & 72.41\% & 0.98 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/recipient4/recipient_pub}  \\
3 & Friends & 69.47\% & 1.02 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/recipient4/recipient_friends}\\
4 & App's Server & 42.28\% & 1.15 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/recipient4/recipient_app}\\
\hline
\end{tabular}
\caption{The overall upset rate for all recipients.}
\label{recipient-table}
\end{center}
\end{table}

{\color {red} TODO: table of top and bottom 10 data types by recipient; refer to appendix; shoutout to when you were lying nervous or stressed work}

### 2015-03-04 tex-inputs/results-type.tex
abstract redo, intro tweaks, and much progress on data type results section

\subsubsection{Data Type}
Based on our data, we observed that the largest effect stemmed from the data type being shared in a scenario. We present various statistical models in section \ref{sec:regression} to support this conclusion. The 10 most and 10 least concerning data types can be seen in Table \ref{top10}.

Regardless of the data recipient or the device, participants were most concerned about photos and videos, especially if they contained embarrassing content, nudity, or financial information. As seen in Table \ref{top10-table}, photos and videos accounted for 5 of the top 10 concerns, {\color {red} and are unanimously considered to be concerning}. Information that could be used to impersonate someone (e.g., usernames/passwords for websites) or invade privacy (photos of someone at home) were also among the most concerning data types.

Also regardless of the other factors, the least concerning data types mostly consisted of information that could be observed through observations of public behavior, such as demographic information (e.g., age, gender, language spoken). {\color {red} As seen in Table \ref{top10-table}, participants had distributed opinions on how risky this information was to be shared. A possibility is that people rated these as unconcerning because of unfamiliarity in what applications would use this data for, or because there does not seem to be any immediate financial, social, or physical consequences from having this information shared.}

@@

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{| r | l | r | l |c |}
\hline
Rank & Data & VUR & sd & Distribution \\
\hline
1 & a video of you unclothed & 95.97\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookavideoofyouunclothedcombined} \\
2 & bank account information & 95.91\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourbankaccountinformationcombined}  \\
3 & social security number & 94.84\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyoursocialsecuritynumbercombined}\\
4 & video of you entering in your PIN & 92.67\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookavideoofyouenteringinyourPINatanATMcombined}\\
5 & a photo of you unclothed & 92.59\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookaphotoofyouunclothedcombined}\\
6 & an incriminating/embarrassing photo of you & 91.39\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookanincriminatingphotoofyoudoingsomethingembarrassingcombined}\\
7 & username and password for websites & 89.55\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourusernameandpasswordforwebsitescombined}\\
8 & credit card information & 88.98\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourcreditcardinformationcombined}\\
9 & an incriminating/embarrassing video of you & 88.41\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookanincriminatingvideoofyoudoingsomethingembarrassingcombined}\\
10 & a random (inward-facing) photo you at home & 87.50\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookphotosofyou(withaninward-facingcamera)athomecombined}\\
 & \vdots & \\
64 & eye movement patterns (for eye tracking) & 40.51\%& number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/scannedyoureyetolearnyoureyepatterns(foreyetracking)combined} \\
65 & when and how much you exercise  & 38.66\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenhowandhowmuchyouexercisecombined}\\
66 & when you are happy or having fun  & 34.75\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenyouwerehappyorhavingfuncombined}\\
67 & which television shows you watch & 30.20\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhattelevisionshowsyouwatchcombined}\\
68 & when you are busy or interruptible  & 29.50\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenyouarebusyorinterruptiblecombined}\\
69 & music from your device  & 28.06\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/copiedanduploadedmusicfromyourdevicecombined}\\
70 & your heart rate & 27.50\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourheartratecombined} \\
71 & your age & 24.29\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyouragecombined}\\
72 & the language you speak & 15.86\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedthelanguageyouwerespeakingcombined}\\
73 & your gender & 15.00\% & number & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourgendercombined}\\ 
\hline
\end{tabular}
\caption{The 10 most and least upsetting data types, across all recipients.}
\label{top10-table}
\end{center}
\end{table}

>>>>

%!TEX root = ../paper.tex

\subsubsection{Data Type}
Based on our data, we observed that the largest effect stemmed from the data type being shared in a scenario. We present various statistical models in section \ref{sec:regression} to support this conclusion. The 10 most and 10 least concerning data types can be seen in Table \ref{top10-table}.

Regardless of the data recipient or the device, participants were most concerned about photos and videos, especially if they contained embarrassing content, nudity, or financial information. As seen in Table \ref{top10-table}, photos and videos accounted for 5 of the top 10 concerns, and are almost unanimously considered to be concerning. Information that could be used to impersonate someone (e.g., usernames/passwords for websites) or invade privacy (photos of someone at home) were also among the most concerning data types.

Also regardless of the other factors, the least concerning data types mostly consisted of information that could be observed through observations of public behavior, such as demographic information (e.g., age, gender, language spoken). As seen in Table \ref{top10-table}, participants had spread distributions in perceptions regarding such information.  A possibility is that people rated these as unconcerning because of unfamiliarity in what applications would use this data for, or because there does not seem to be any immediate financial, social, or physical consequences from having this information shared.

@@

\begin{table*}[t]
\begin{center}
\small
\begin{tabular}{| r | l | r | r |c |}
\hline
Rank & Data & VUR & $\sigma$ & Distribution \\
\hline
1 & video of you unclothed & 95.97\% & 0.31 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookavideoofyouunclothedcombined} \\
2 & bank account information & 95.91\% & 0.35 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourbankaccountinformationcombined}  \\
3 & social security number & 94.84\% & 0.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyoursocialsecuritynumbercombined}\\
4 & video entering in a PIN at an ATM & 92.67\% & 0.47 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookavideoofyouenteringinyourPINatanATMcombined}\\
5 & photo of you unclothed & 92.59\% & 0.46 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookaphotoofyouunclothedcombined}\\
6 & photo of you that is very embarrassing & 91.39\% & 0.55 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookanincriminatingphotoofyoudoingsomethingembarrassingcombined}\\
7 & username and password for websites & 89.55\% & 0.62 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourusernameandpasswordforwebsitescombined}\\
8 & credit card information & 88.98\% & 0.56 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourcreditcardinformationcombined}\\
9 & video of you that is very embarrassing & 88.41\% & 0.53 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookanincriminatingvideoofyoudoingsomethingembarrassingcombined}\\
10 & photo of you at home & 87.50\% & 0.60 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/tookphotosofyou(withaninward-facingcamera)athomecombined}\\
 & \vdots & & & \\
64 & eye patterns (for eye tracking) & 40.51\%& 1.27 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/scannedyoureyetolearnyoureyepatterns(foreyetracking)combined} \\
65 & exercise patterns  & 38.66\% & 1.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenhowandhowmuchyouexercisecombined}\\
66 & when you are happy or having fun  & 34.75\% & 1.27 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenyouwerehappyorhavingfuncombined}\\
67 & television shows watched & 30.20\% & 1.40 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhattelevisionshowsyouwatchcombined}\\
68 & when you are busy or interruptible  & 29.50\% & 1.26 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedwhenyouarebusyorinterruptiblecombined}\\
69 & music on device  & 28.06\% & 1.43 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/copiedanduploadedmusicfromyourdevicecombined}\\
70 & your heart rate & 27.50\% & 1.40 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourheartratecombined} \\
71 & age & 24.29\% & 1.43 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyouragecombined}\\
72 & language spoken & 15.86\% & 1.49 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedthelanguageyouwerespeakingcombined}\\
73 & gender & 15.00\% & 1.45 & \includegraphics[width = 2cm, height = 0.5cm]{tex-inputs/data10/learnedyourgendercombined}\\ 
\hline
\end{tabular}
\caption{The 10 most and least upsetting data types, across all recipients. For the complete list of all data types across all recipients, see Appendix [?].}
\label{top10-table}
\end{center}
\end{table*}

The data types examined span several existing and future use cases. We do not believe they are comprehensive in coverage of how wearable devices might capture data nor for all possible future use cases of the data. Therefore, we coded each data type in two ways: firstly in terms in the data and how it was captured (e.g., video, audio, text, etc.), and  secondly in terms of the type of risks presented to the user. This examines why participants viewed certain data collected as more/less risky than others, letting us  generalize our results. Two researchers agreed on a codebook and independently coded each of the 72 data types\footnote{We excluded the data types that did not feature a data recipient.}

The data captured fell into the following six possible categories:

\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
\item Photo
\item Video
\item Audio
\item Behavioral Information
\item Biometric Information
\item Demographic Information
\end{enumerate}

The associated risks data type fell into the following five categories:

\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
\item {\bf Financial:} The risk involves the loss of money or property.
\item {\bf Image:} The risk involves loss of control over one's self-image (e.g., publicizing something embarrassing).
\item {\bf Medical:} The risk involves disclosure of medical information.
\item {\bf Physical:} The risk involves physical harm to the user.
\item {\bf Relationships:} The risk involves damage to the user's inter-personal relationships.
\end{enumerate}

After independently coding all data types with respect to these categories, the researchers met to resolve any disagreements, such that the resulting codings reflected unanimity. Prior to this resolution, there was  83\% agreement. Cohen's $\kappa$ is 0.81 for the data capture and 0.75 for the associated risks, both indicating ``excellent'' agreement~\cite{Fleiss2003}.
